---
title: "Analysis"
output: html_document
authors: "Anna Cameron, Enya Tan, Mubarak Ganiyu, Sriram Kannan"
---

### Loading packages

```{r}
library(tidyverse)
library(janitor)
library(readr)
library(assertr)
library(ggradar)
library(tidytext)
library(tidyr)
```

### Importing data

```{r}
music <- read_csv("~/Projects/Vandy Assignments/EDA-project/clean-data/tcc_ceds_music_cleaned.csv")
```

## Analysis

### Which Genre of music has been produced the most?

```{r}
musiccor <- music %>% select(danceability:Length_norm, -topic)
corrplot(cor(musiccor), method = "square", tl.cex = 0.9)
```


```{r}
music_bar <- music %>% 
  count(genre) %>% # counting number of genres
  na.omit() %>% # removing NA values
  ungroup() %>% 
  mutate(prop = (n/sum(n))*100) %>% # calculating proportions of accidents
  mutate(label = as.character(prop)) 
music_bar$label <- substr(music_bar$label, 0, 4) # converting proportions into characters for labelling
ggplot(music_bar) + ## Plotting graphic
  geom_bar(aes(x = "", y = prop, fill = genre), stat= "identity", width = 1) +
  labs(title = "Tracks by Genre", fill = "Genre") +
  coord_polar("y", start=0) +
  theme_void() +
  theme(plot.title = element_text(hjust=0.5))
```

### Correlational Analysis of Musical Features across Genres

#### Creating function for calculating correlation

```{r}
correlation <- function(vec, x, y) {
  val <- c()
  for (i in seq_along(vec)) {
    gen <- music %>% 
      filter(genre == vec[i])
    val[i] <- cor(gen[,x], gen[,y], use="everything",
        method="pearson")
  }
  df <- data.frame(genre = vec, corr = val, abs_corr = abs(val))
  return(df)
}
```

#### Valence and Danceability correlation across different genres

```{r, fig.width = 10, fig.height = 8}
gen_vec <- unique(music$genre)
new_df <- correlation(gen_vec, "valence", "danceability")
new_df %>% 
  arrange(desc(corr))

ggplot(new_df) +
  geom_bar(aes(x = reorder(genre, -abs_corr), y = abs_corr), stat = "identity",
   fill = "skyblue", alpha = 0.7) +
  labs(x = "Genre", y = "Valence and Danceability Correlation",
       title = "Valence and Danceability Correlation across different Genres") +
  theme_light() +
  theme(plot.title = element_text(hjust =0.5))
  
```

#### Age and Acousticness correlation across different genres

```{r, fig.width = 10, fig.height = 8}
new_df <- correlation(gen_vec, "age", "acousticness")
new_df %>% 
  arrange(desc(corr))

ggplot(new_df) +
  geom_bar(aes(x = reorder(genre, -abs_corr), y = abs_corr), stat = "identity",
   fill = "skyblue", alpha = 0.7) +
  labs(x = "Genre", y = "Age and Acousticness Correlation",
       title = "Age and Acousticness Correlation across different Genres") +
  theme_light() +
  theme(plot.title = element_text(hjust =0.5))
```

#### Age and Energy correlation across different genres

```{r, fig.width = 10, fig.height = 8}
new_df <- correlation(gen_vec, "age", "energy")
new_df %>% 
  arrange(desc(corr))

ggplot(new_df) +
  geom_bar(aes(x = reorder(genre, -abs_corr), y = abs_corr), stat = "identity",
   fill = "skyblue", alpha = 0.7) +
  labs(x = "Genre", y = "Age and Energy Correlation",
       title = "Age and Energy Correlation across different Genres") +
  theme_light() +
  theme(plot.title = element_text(hjust =0.5))
```

#### Age and Loudness correlation across different genres

```{r, fig.width = 10, fig.height = 8}
new_df <- correlation(gen_vec, "age", "loudness")
new_df %>% 
  arrange(desc(corr))

ggplot(new_df) +
  geom_bar(aes(x = reorder(genre, -abs_corr), y = abs_corr), stat = "identity",
   fill = "skyblue", alpha = 0.7) +
  labs(x = "Genre", y = "Age and Loudness Correlation",
       title = "Age and Loudness Correlation across different Genres") +
  theme_light() +
  theme(plot.title = element_text(hjust =0.5))
```

#### Energy and Loudness correlation across different genres

```{r, fig.width = 10, fig.height = 8}
gen_vec <- unique(music$genre)
new_df <- correlation(gen_vec, "energy", "loudness")
new_df %>% 
  arrange(desc(corr))

ggplot(new_df) +
  geom_bar(aes(x = reorder(genre, -abs_corr), y = abs_corr), stat = "identity",
   fill = "skyblue", alpha = 0.7) +
  labs(x = "Genre", y = "Energy and Loudness Correlation",
       title = "Energy and Loudness Correlation across different Genres") +
  theme_light() +
  theme(plot.title = element_text(hjust =0.5))
  
```

### Exploring how genres differ on average based on different features

#### Audio parameter differences across genres

```{r, fig.width = 10, fig.height = 8}
radar2 <- music %>% select(genre, energy, loudness, valence, danceability, acousticness, instrumentalness, Length_norm)

radar2g <- radar2 %>% group_by(genre) %>% summarize(Energy = mean(energy), Loudness = mean(loudness), Valence = mean(valence), Danceability = mean(danceability), Acoustincness = mean(acousticness), Instrumentalness = mean(instrumentalness), Length = mean(Length_norm))

ggradar(radar2g, values.radar = c('', '', ''), legend.title = "Genre", legend.text.size = 10, axis.label.size = 3)
```



```{r}
music_new <- music %>% select(genre, track_name, lyrics) %>% group_by(genre) %>% unnest_tokens(word, lyrics)

sentiments <- get_sentiments("nrc")

df_sentiments <- music_new %>% left_join(sentiments)

df_sentiments_filtered <- df_sentiments %>% filter(!is.na(sentiment)) %>% group_by(genre, sentiment) %>% summarize(count = n())

df_sentiments_filtered$count <- normalize(df_sentiments_filtered$count)

df_sentiments_filtered
```

#### Sentimental differences across genres

```{r, fig.width = 10, fig.height = 8}
(genresentiments <- df_sentiments_filtered %>% pivot_wider(names_from = sentiment, values_from = count))
ggradar(genresentiments, values.radar = c('', '', ''), legend.title = "Genre", legend.text.size =8, axis.label.size = 3)
```


### Genre changes on average according to different musical features

#### Change in Loudness over time (1950-2019) across genre

```{r, fig.width = 12, fig.height = 9}
music_agg <- music %>% 
  group_by(genre, release_date) %>% 
  summarize(mean = mean(loudness))
ggplot(music_agg) + 
  geom_line(aes(x = release_date, y = mean, color = genre)) +
  geom_point(aes(x = release_date, y = mean, color = genre)) +
  labs(title="Change in Loudness over Time (1950-2019) across Genre",
       x="Year",
       y= "Loudness") +
  scale_x_continuous(breaks= seq(1950, 2020, 25)) +
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Change in Acousticness over time (1950-2019) across genre

```{r, fig.width = 12, fig.height = 9}
music_agg <- music %>% 
  group_by(genre, release_date) %>% 
  summarize(mean = mean(acousticness))
ggplot(music_agg) + 
  geom_line(aes(x = release_date, y = mean, color = genre)) +
  geom_point(aes(x = release_date, y = mean, color = genre)) +
  labs(title="Change in Acousticness over Time (1950-2019) across Genre",
       x="Year",
       y= "Energy") +
  scale_x_continuous(breaks= seq(1950, 2020, 25)) +
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Change in Energy over time (1950-2019) across genre

```{r, fig.width = 12, fig.height = 9}
music_agg <- music %>% 
  group_by(genre, release_date) %>% 
  summarize(mean = mean(energy))
ggplot(music_agg) + 
  geom_line(aes(x = release_date, y = mean, color = genre)) +
  geom_point(aes(x = release_date, y = mean, color = genre)) +
  labs(title="Change in Energy over Time (1950-2019) across Genre",
       x="Year",
       y= "Energy") +
  scale_x_continuous(breaks= seq(1950, 2020, 25)) +
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5))
```


### How the genres vary with regards to the topics affiliated to the song

```{r, fig.width = 12, fig.height = 9}
music_count <- music %>% 
  group_by(genre) %>% 
  summarize(count = n())
music_count
ggplot(music) +
  geom_bar(aes(x = genre, fill = topic), position="fill",
           color = "rosybrown", alpha = 0.7) +
  labs(x = "Genre", title = "How Genres differ in terms of Topic", y = "proportion") +
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5))
```

### 

```{r}
ggplot(music) +
  geom_histogram(aes(x = len)) + labs(title = "Distribution of Length of Songs", x = "Length of Songs", y = "Frequency") + theme(plot.title = element_text(hjust = 0.5)) 
```

```{r, fig.width = 12, fig.height = 9}
ggplot(music) +
  geom_boxplot(aes(y = len, x = genre, color = genre), fill = "gray45") +
  labs(x = "Genre", y = "Length", title = "Distributions of Length of Songs by Genre") +
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5))
```
